Supporting information: Copulas and their potential for ecology

### 2019-12-18 SupportingInformation.Rmd
changed distribution function to cdf

Let $F$ be the bivariate distribution function
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$ assumed to be 
continuous and strictly monotonic. We show the copula associated with
$(X,Y)$, which exists and is unique by Sklar's theorem, is, in fact, 
the distribution function
of the random variable $(U,V)=(F_X(X),F_Y(Y))$. Let $C$ denote the distribution
function of $(U,V)$, which has uniform margins. By the uniqueness
statement of Sklar's theorem, it suffices to show that $F(x,y)=C(F_X(x),F_Y(y))$
for all $(x,y)$ in the Euclidean plane. But 
$C(F_X(x),F_Y(y)) = P[U \leq F_X(x), V \leq F_Y(y)]
= P[F_X(X) \leq F_X(x), F_Y(Y) \leq F_Y(y)]
= P[X \leq x, Y \leq y]
= F(x,y),$
where the third equality follows from the strict monotonicity assumption. In fact, 
that equality should hold even with non-strict monotonicity, so a stronger 
statement is possible (see below for more on this topic). The continuity assumption 
for $F_X$ and $F_Y$ is necessary, however: without it the random variables 
$U$ and $V$ are not uniformly distributed, so their joint distribution function 
is not even a copula.

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate distribution functions, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a distribution function.
We here show that if $G_A$ and $G_B$ are continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To get to the desired result, note that the distribution function of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
$(U,V)$. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.

>>>>

Let $F$ be the cdf
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$ assumed to be 
continuous and strictly monotonic. We show the copula associated with
$(X,Y)$, which exists and is unique by Sklar's theorem, is, in fact, 
the bivariate cdf
of the random variable $(U,V)=(F_X(X),F_Y(Y))$. Let $C$ denote the distribution
function of $(U,V)$, which has uniform margins. By the uniqueness
statement of Sklar's theorem, it suffices to show that $F(x,y)=C(F_X(x),F_Y(y))$
for all $(x,y)$ in the Euclidean plane. But 
$C(F_X(x),F_Y(y)) = P[U \leq F_X(x), V \leq F_Y(y)]
= P[F_X(X) \leq F_X(x), F_Y(Y) \leq F_Y(y)]
= P[X \leq x, Y \leq y]
= F(x,y),$
where the third equality follows from the strict monotonicity assumption. In fact, 
that equality should hold even with non-strict monotonicity, so a stronger 
statement is possible (see below for more on this topic). The continuity assumption 
for $F_X$ and $F_Y$ is necessary, however: without it the random variables 
$U$ and $V$ are not uniformly distributed, so their joint cdf 
is not even a copula.

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate cdfs, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a cdf.
We here show that if $G_A$ and $G_B$ are continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the cdf of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To get to the desired result, note that the cdf of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the cdf of 
$(U,V)$. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.
